### 2024-09-25 proposal/project_proposal.tex
started the data section

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% for space filling
\usepackage{lipsum}
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data

\title{Proposal: An Analysis of Hockey's Most Controversial Statistic}
\author{Melanie Desroches\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}
In hockey, a very common statistic that is reported is the plus/minus statistic. Plus/minus is calculated
by taking all the even strength goals that were scored by their team while they were on the ice and subtracting
it by the total number of goals that were scored by the other team while they were on the ice. The idea behind 
the number is that it will show if a player has an overall positive or negative contribution when they are playing. 
However, the plus/minus has a lot of drawbacks. For example, in ice hockey, players tend to play on "lines", meaning 
the same three forwards and the same two defencemen tend to play at the same time. Because of this, plus/minus is heavily 
influenced by who they play with. Some would argue that the plus/minus is more of a team statistic rather than an 
individual statistics. Many of the criticisms of this statistic comes from the fact that there are a lot of confounding 
variables that influence how the plus/minus is calculated. Because of these criticisms, alternatives to the plus/minus
statistic have been developed, such as COSRI or Fenwick.

The goal of this paper is to evaluate the effectiveness of the plus/minus statistic. Some of the information
I am trying to determine is if the plus/minus statistic can be used to determine how good a player is both on
offense and defense. I also want to see if the plus/minus has any validity being used as an individual
statistic or as a team statistic. Based on my observations, I will compare the plus/minus statistic to other
statistics that have a similar purpose such as CORSI or Fenwick.

\paragraph{Specific Aims}
Formulate your research question;
translate your research question into statistical/data science questions

Research questions:
Is the plus/minus statistic an effective way of evaluating a players offensive or defensive
contribution?
Is the plus/minus statistic an individual or team statistic?
Is the plus/minus statistic a better evaluation of a player than other contribution statistics
such as CORSI or Fenwick?

\paragraph{Data}
Hopefully, you have identified the data needed for your project. Give a
description about it.

I am using data from the National Hockey League (NHL). The data is collected from a website called
Hockey Reference. I am particularly interested in the advanced hockey statistics that are collected
for each player in the NHL.

\paragraph{Research Design and Methods}
What design or methods will you use?

Since the plus/minus statistic tends to deal with collinearity, a ridge regression might be more
appropriate for analysis.

\paragraph{Discussion}
What are the most challenge parts of the task?
What are the limitations of your work? What is your fall-back plan if
something unexpected happens?

\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}